Type Refinements for the Working Class

### 2015-11-22 paper.tex
Define subrefinement relation, equality of refinements

\begin{definition}
  To know $\Refines{\Upsilon}{\phi}{\tau}$ (presupposing $\IsSymCtx{\Upsilon}$
  and $\IsSort{\tau}$) is to know, for any $\Of{\rho}{\Upsilon\inj\Upsilon'}$,
  what it means for any $\Member{M,N}{\Exprs_\tau(\Upsilon')}$ to be equated by
  $\phi$ (written $\RefEquate{\phi}{\rho}{M}{N}$), such that
  $\RefEquate{\phi}{\rho}{\cdot}{\cdot}$ is a partial equivalence relation on
  $\Exprs_\tau(\Upsilon')$. Morever, that for any
  $\Of{\rho'}{\Upsilon'\inj\Upsilon''}$, from $\RefEquate{\phi}{\rho}{M}{N}$ we
  may conclude $\RefEquate{\phi}{\rho'\circ\rho}{M\rho'}{N\rho'}$.
\end{definition}

@@

@@

%% \subsection{Parametric Refinement}
%% We will require two judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$,
%% which means that $\Phi$ is a symbol context which refines the symbol context
%% $\Upsilon$ (presupposing $\IsSymCtx{\Upsilon}$), and parametric refinement
%% $\Refines{\Phi}{\phi}{\tau}$ (which shall presuppose
%% $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
%% \newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
%% \newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
%% The first judgment we define inductively:
%%   \infer[\RefinesCtxNil]{
%%     \RefinesCtx{\cdot}{\cdot}
%%   \infer[\RefinesCtxSnoc]{
%%     \RefinesCtx{\Phi,u:\phi}{\Upsilon,u:\tau}
%%     \RefinesCtx{\Phi}{\Upsilon} &
%%     \Refines{\Phi}{\phi}{\tau}
%% The second judgment is defined coinductively via the following meaning
%% explanation:
%% \begin{definition}[Parametric Refinement]
%%   To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
%%   $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
%%   $\Of{\varrho}{\Upsilon\inj\OMode{\Upsilon'}}$, and for any $M, N$,
%%   what it means for $M$ and $N$ to be identified under $\phi$, supposing
%%   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
%%   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
%%   $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
%%   $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
%%   $\IMode{M}\sim^\rho_\phi\IMode{O}$.
%% \end{definition}
%% In other words, to know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
%% $\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
%% $\Upsilon'$, which partial equivalence relation on the closed $\tau$-sorted
%% $\Upsilon'$-terms $\phi$ denotes. At this point, it should be remarked that
%% this is very similar to the semantical explanation of typehood given in
%% \cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
%% setting, and that we have fibred the entire apparatus over collections
%% $\Upsilon$ of symbols.
%% \todo[inline]{%
%%   define maps of refined contexts, then change weakening lemma to
%%   monotonicity, which is stronger and more useful.
%% \begin{thm}[Weakening]
%%   If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
%%   $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
%% \end{thm}
%% \begin{proof}
%%   This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
%%   $\Upsilon\inj\Upsilon,u:\sigma$.
%% \end{proof}
%% When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ \emph{globally}
%% refines $\tau$ and may simply write $\GRefines{\phi}{\tau}$. By weakening, if
%% $\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
%% $\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.
%% \subsubsection{Ordering and Equality of Parametric Refinement}
%% We will write $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ to mean that $\phi$ is a
%% \emph{subrefinement} of $\psi$, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
%% $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$:
%% \begin{definition}
%%   To know $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
%%   $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
%%     \GenJ{
%%       \MV{m},\MV{n}:\tau
%%         \IMode{M}\sim^\rho_\psi\IMode{N}
%%         \IMode{M}\sim^\rho_\phi\IMode{N}
%% \end{definition}
%% Two refinements are equal when they denote the same PER. That is, we have
%% $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ just when both
%% $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ and $\SubRefines{\Phi}{\psi}{\phi}{\tau}$.

@@

Now, we can extend $R_\tau$ to a new relation $[R_\tau]$ on closed expressions
which respects a single ``layer'' of computation. We will say
$\SBinRel{[R_\tau]}{\Upsilon}{M}{N}$ when, supposing
$\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}}$ such that
$\IMode{\Upsilon}\Vdash_{\IMode{\CanOperators}}\IMode{\vartheta}:\OMode{\MkArity{\vec{v}}{\tau}}$,
for any $\Of{\rho}{\Upsilon\inj\Upsilon'}$, we have
$\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and
$\SBinRel{R_{\vec{v}}}{\Upsilon'}{\vec{E}\rho}{\vec{F}\rho}$.

>>>>

\begin{definition}
  To know $\Refines{\Upsilon}{\phi}{\tau}$ (presupposing $\IsSymCtx{\Upsilon}$
  and $\IsSort{\tau}$) is to know, for any
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd},
  what it means for any $\Member{M,N}{\Exprs_\tau(\Upsilon')}$ to be equated by
  $\phi$ (written $\RefEquate{\phi}{\rho}{M}{N}$), such that
  $\RefEquate{\phi}{\rho}{\cdot}{\cdot}$ is a partial equivalence relation on
  $\Exprs_\tau(\Upsilon')$. Morever, that for any
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon'}\arrow[r,hook,"\IMode{\rho'}"] &\IMode{\Upsilon''}
  \end{tikzcd},
  from $\RefEquate{\phi}{\rho}{M}{N}$ we may conclude
  $\RefEquate{\phi}{\rho'\circ\rho}{M\rho'}{N\rho'}$.
\end{definition}

@@

\subsubsection{Order and Equality of Parametric Refinements}

We will write $\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ to mean that $\phi$ is
a \emph{subrefinement} of $\psi$.

\begin{definition}

  To know $\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ (presupposing
  $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$) , is
  to know, for any renamings
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r, hook, "\IMode{\rho}"]
    & \IMode{\Upsilon'} \arrow[r, hook, "\IMode{\rho'}"]
    & \IMode{\Upsilon''}
  \end{tikzcd},
  that from $\RefEquate{\phi}{\rho}{M}{N}$ you can conclude
  $\RefEquate{\psi}{\rho'\circ\rho}{M\rho'}{N\rho'}$ for any
  $\Member{M,N}{\Exprs_\tau(\Upsilon')}$.

Two refinements are equal when they denote the same PER. That is, we have
$\EqRefines{\Upsilon}{\phi}{\psi}{\tau}$ just when both
$\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ and
$\SubRefines{\Upsilon}{\psi}{\phi}{\tau}$.

@@

@@

Now, we can extend $R_\tau$ to a new relation $[R_\tau]$ on closed expressions
which respects a single ``layer'' of computation. We will say
$\SBinRel{[R_\tau]}{\Upsilon}{M}{N}$ when, supposing
$\Eval{\Upsilon}{\tau}{M}{\App{\vartheta}{\Vec{E}}}$ such that
$\IMode{\Upsilon}\Vdash_{\IMode{\CanOperators}}\IMode{\vartheta}:\OMode{\MkArity{\vec{v}}{\tau}}$,
for any
\begin{tikzcd}[cramped]
  \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
\end{tikzcd},
we have $\Eval{\Upsilon'}{\tau}{N\rho}{\App{\vartheta\rho}{\Vec{F}\rho}}$ and
$\SBinRel{R_{\vec{v}}}{\Upsilon'}{\vec{E}\rho}{\vec{F}\rho}$.
